---
title: "My project"
author: "Casey Zhang"
date: "2024-10-10"
output: "github_document"
---
# Load packages and dataset
```{r}
library(haven) #to load datasets (.csv, .hav)
library(dplyr) #to recode variables
library(ggplot2) #density plots, etc.
library(tidyr)

library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)
library(ggsci)
library(see)

library(tidyverse)
library(bruceR)
library(haven)
#when you make changes, press knit, commit to main - changes will be reflected in github desktop & online


load("/Users/caseyzhang/Desktop/ICPSR_36850/DS0001/36850-0001-Data.rda")
```

# Narrow Down Data
```{r}
Project_data <- da36850.0001 %>%
  select(DEMGEN, GRIT01, GRIT02, GRIT03, GRIT04, GRIT05, GRIT06, GRIT07, GRIT08, TSCS01, TSCS02, TSCS03, TSCS04, TSCS05, TSCS06, TSCS07, TSCS08, TSCS09, DEMRA010)

#Missing values: listwise deletion
Project_data<- drop_na(Project_data)

```

# Recode Variables
```{r}

#Gender: Change Numeric Values to Characters

Project_data %>% count(DEMGEN)

Project_data<- Project_data%>%
  mutate_at(c('DEMGEN'),funs(str_replace(., "\\(1\\) Female", "1")))

Project_data<- Project_data%>%
  mutate_at(c('DEMGEN'),funs(str_replace(., "\\(0\\) Male", "0")))

#Re-code reverse coded items: grit(1,3,5), self control(4,6,9)
Project_data$GRIT01_R <- 5 - Project_data$GRIT01
Project_data$GRIT03_R <- 5 - Project_data$GRIT03
Project_data$GRIT05_R <- 5 - Project_data$GRIT05
Project_data$TSCS04_R <- 5 - Project_data$TSCS04
Project_data$TSCS06_R <- 5 - Project_data$TSCS06
Project_data$TSCS09_R <- 5 - Project_data$TSCS09

#Race: Change Numeric Values to Characters
Project_data %>% count(DEMRA010)

Project_data<- Project_data%>%
  mutate(DEMRA010 = str_replace_all(DEMRA010, c("\\(1\\) Yes" = "1", "\\(0\\) No" = "0")))

```
# Create Composites (grit, self control)
```{r}

Project_data <- Project_data %>%
  mutate(GRIT = rowMeans(cbind(GRIT01_R, GRIT02, GRIT03_R, GRIT04, GRIT05_R, GRIT06, GRIT07, GRIT08)))

Project_data <- Project_data %>%
  mutate(TSCS = rowMeans(cbind(TSCS01, TSCS02, TSCS03, TSCS04_R, TSCS05, TSCS06, TSCS07, TSCS08, TSCS09_R)))

```

# Correlation between self-control and grit
```{r}

Project_data_cor <- Project_data %>%
  select(GRIT, TSCS)

Corr(Project_data_cor)

# Ensure Project_data_cor is correctly created
Project_data_cor <- Project_data %>%
  select(GRIT, TSCS)

# Calculate the correlation coefficient
correlation <- cor(Project_data_cor$GRIT, Project_data_cor$TSCS, use = "complete.obs")

# Print the correlation to verify
print(paste("Correlation coefficient:", round(correlation, 2)))

# Plot the scatter plot
plot(Project_data_cor$TSCS, Project_data_cor$GRIT,
     main = paste("Correlation Between Self-control and Grit (r =", round(correlation, 2), ")"),
     xlab = "Self-control", ylab = "Grit",
     pch = 19, col = "blue")

# Add trend line
abline(lm(TSCS ~ GRIT, data = Project_data_cor), col = "red")


```

# Test whether the correlation is moderated by gender
```{r}

PROCESS(Project_data, y = "GRIT", x = "TSCS", mods = c("DEMGEN"))

```

#Visualize the correlation overall & correlation grouped by gender
```{r}
ggplot(Project_data, aes(x = GRIT, y = TSCS)) + geom_point() + geom_smooth(method = lm) + theme_bruce() 
ggplot(Project_data, aes(x = GRIT, y = TSCS)) + geom_point() + geom_smooth(method = lm) + theme_bruce() +facet_wrap(~DEMGEN)



```

# Exploratory Question: Test whether the correlation is moderated by race & visualize the correlation grouped by race
```{r}

PROCESS(Project_data, y = "GRIT", x = "TSCS", mods = c("DEMRA010"))
ggplot(Project_data, aes(x = GRIT, y = TSCS)) + geom_point() + geom_smooth(method = lm) + theme_bruce() +facet_wrap(~DEMRA010)

```

#Reliability
Variable 1: Self-control, TSCS (9 items, R=4,6,9)
Variable 2: Grit, GRIT (8 items, R=1,3,5)

The cronbach's alpha for self-control is 0.697 and for grit it is 0.657, both of which are considered acceptable.
```{r}

Alpha(Project_data, "TSCS0", 1:9) 

Alpha(Project_data, "GRIT0", 1:8)

```
# Exploratory Factor Analysis
For self-control, it detects 2 factors because 2 factors have Eigenvalues above 1, there are 2 points before the flattened point of the scree plot, and there are 2 actual dots above the simulated scree plot. However, the detection of 2 factors was a result of the reverse-coded items so 1 factor was forced. The range of factor loadings and percentage of variance explained for factor 1 is 0.181-0.629 and 31.092%. Overall, the measurement validity is unacceptable to good based on factor 1.

For grit, it detects 2 factors because 2 factors have Eigenvalues above 1, there are 2 points before the flattened point of the scree plot, and there are 2 actual dots above the simulated scree plot. However, again, the detection of 2 factors was a result of the reverse-coded items so 1 factor was forced. The range of factor loadings and percentage of variance explained for factor 1 is 0.258-0.720 and 31.187%. Overall, the measurament validity is unacceptable to good based on factor 1.

These 
```{r}

EFA(Project_data, "TSCS0", 1:9, method = "pa", plot.scree = TRUE, nfactors = 1)
EFA(Project_data, "GRIT0", 1:8, method = "pa", plot.scree = TRUE, nfactors = 1)
```


