---
title: "My project"
author: "Casey Zhang"
date: "2024-10-10"
output: "github_document"
---
# Load packages and dataset
```{r}
library(haven) #to load datasets (.csv, .hav)
library(dplyr) #to recode variables
library(ggplot2) #density plots, etc.
library(tidyr)
#when you make changes, press knit, commit to main - changes will be reflected in github desktop & online


load("/Users/caseyzhang/Desktop/ICPSR_36850/DS0001/36850-0001-Data.rda")
```

# Narrow Down Data
```{r}
Project_data <- da36850.0001 %>%
  select(DEMGEN, GRIT01, GRIT02, GRIT03, GRIT04, GRIT05, GRIT06, GRIT07, GRIT08, TSCS01, TSCS02, TSCS03, TSCS04, TSCS05, TSCS06, TSCS07, TSCS08, TSCS09, DEMRA010)

#Missing values: listwise deletion
Project_data<- drop_na(Project_data)

```

# Recode Variables
```{r}

#Gender: Change Numeric Values to Characters

Project_data %>% count(DEMGEN)

Project_data<- Project_data%>%
  mutate_at(c('DEMGEN'),funs(str_replace(., "\\(1\\) Female", "1")))

Project_data<- Project_data%>%
  mutate_at(c('DEMGEN'),funs(str_replace(., "\\(0\\) Male", "0")))

#Re-code reverse coded items: grit(1,3,5), self control(4,6,9)
Project_data$GRIT01_R <- 5 - Project_data$GRIT01
Project_data$GRIT03_R <- 5 - Project_data$GRIT03
Project_data$GRIT05_R <- 5 - Project_data$GRIT05
Project_data$TSCS04_R <- 5 - Project_data$TSCS04
Project_data$TSCS06_R <- 5 - Project_data$TSCS06
Project_data$TSCS09_R <- 5 - Project_data$TSCS09

#Race: Change Numeric Values to Characters
Project_data %>% count(DEMRA010)

Project_data<- Project_data%>%
  mutate(DEMRA010 = str_replace_all(DEMRA010, c("\\(1\\) Yes" = "1", "\\(0\\) No" = "0")))

```
# Create Composites (grit, self control)
```{r}

Project_data <- Project_data %>%
  mutate(GRIT = rowMeans(cbind(GRIT01_R, GRIT02, GRIT03_R, GRIT04, GRIT05_R, GRIT06, GRIT07, GRIT08)))

Project_data <- Project_data %>%
  mutate(TSCS = rowMeans(cbind(TSCS01, TSCS02, TSCS03, TSCS04_R, TSCS05, TSCS06, TSCS07, TSCS08, TSCS09_R)))

```

# Correlation between self-control and grit
```{r}

Project_data_cor <- Project_data %>%
  select(GRIT, TSCS)

Corr(Project_data_cor)

```

# Test whether the correlation is moderated by gender
```{r}

PROCESS(Project_data, y = "GRIT", x = "TSCS", mods = c("DEMGEN"))

```

#Visualize the correlation overall & correlation grouped by gender
```{r}
ggplot(Project_data, aes(x = GRIT, y = TSCS)) + geom_point() + geom_smooth(method = lm) + theme_bruce() 
ggplot(Project_data, aes(x = GRIT, y = TSCS)) + geom_point() + geom_smooth(method = lm) + theme_bruce() +facet_wrap(~DEMGEN)

```

# Exploratory Question: Test whether the correlation is moderated by race & visualize the correlation grouped by race
```{r}

PROCESS(Project_data, y = "GRIT", x = "TSCS", mods = c("DEMRA010"))
ggplot(Project_data, aes(x = GRIT, y = TSCS)) + geom_point() + geom_smooth(method = lm) + theme_bruce() +facet_wrap(~DEMRA010)

```

#Reliability
Note: I tried the way we learned in class (code shown below) and it didn't work with my dataset for some reason so I used different methods.

Variable 1: Self-control, TSCS (9 items, R=4,6,9)
Variable 2: Grit, GRIT (8 items, R=1,3,5)

The cronbach's alpha for self-control is 0.33 and for grit it is 0.02, both of which are poor.
```{r}

#Self-control
colnames(Project_data)
selected_columns <- Project_data[, c("TSCS01", "TSCS02", "TSCS03", "TSCS04_R", "TSCS05", "TSCS06_R", "TSCS07", "TSCS08", "TSCS09_R")]
library(psych)
alpha_result <- alpha(selected_columns)
print(alpha_result)

#Grit
colnames(Project_data)
selected_columns <- Project_data[, c("GRIT01_R", "GRIT02", "GRIT03_R", "GRIT04", "GRIT05_R", "GRIT06", "GRIT07", "GRIT08", "GRIT05")]
library(psych)
alpha_result <- alpha(selected_columns)
print(alpha_result)

#Didn't work
Alpha(Project_data, "TSCS", c("1", "2", "3", "4_R", "5", "6_R" , "7" , "8" , "9_R"))
Alpha(Project_data, "TSCS", 1:9, rev = 4, 6, 9)

Alpha(Project_data, "GRIT", c("1_R", "2", "3_R", "4_R", "5_R", "6" , "7" , "8" , "9"))
Alpha(Project_data, constrains("GRIT"), 1:8, rev = 1,3,5)
```
# Exploratory Factor Analysis
For self-control, two factors have Eigenvalues above 1, there are 2 points before the flattened point of the scree plot, and there are 2 actual dots above the simulated scree plot, suggesting there are 2 factors. The range of factor loadings and percentage of variance explained is -0.207-0.569 and 31.092% for factor 1 and -0.460-0.671 and 15.225% for factor 2. Overall, the measurament validity is unacceptable to good based on factor 1 and unacceptable to acceptable based on factor 2.

For grit, two factors have Eigenvalues above 1, there are 2 points before the flattened point of the scree plot, and there are 2 actual dots above the simulated scree plot, suggesting there are 2 factors. The range of factor loadings and percentage of variance explained is -0.143-0.820 and 31.187% for factor 1 and -0.198-0.515 and 18.667% for factor 2. Overall, the measurament validity is unacceptable to good based on factor 1 and unacceptable to acceptable based on factor 2.
```{r}

#Self-control
tscs_columns <- c("TSCS01", "TSCS02", "TSCS03", "TSCS04", "TSCS05", "TSCS06", "TSCS07", "TSCS08", "TSCS09")
EFA(Project_data, "TSCS", vars = tscs_columns, rev = c("TSCS04", "TSCS06", "TSCS09"), 
    method = "pa", plot.scree = TRUE, nfactors = "parallel")

#Grit
grit_columns <- c("GRIT01", "GRIT02", "GRIT03", "GRIT04", "GRIT05", "GRIT06", "GRIT07", "GRIT08")
EFA(Project_data, "GRIT", vars = grit_columns, rev = c("GRIT01", "GRIT03", "GRIT05"), 
    method = "pa", plot.scree = TRUE, nfactors = "parallel")

#Didn't work
EFA(Project_data, "TSCS", 1:9, rev = 4,6,9, method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
EFA(Project_data, "GRIT", 1:8, rev = 1,3,5, method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
```


